---
title: "Practice_Facilit_arid"
output: html_document
date: "2024-09-01"
---

##The aim of the following practice is that you answer some relevant questions that can be aproached using the data provided in RecruitNet to build a Facilitation network, using as an example the study site: Datcamaquis


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##########################################################################################################
# 1. Load the functions and that data
##########################################################################################################


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

##############################################################################################################
##2. extract from RecruitNet database the information from the study site:
##############################################################################################################

```{r}


```


##################################################################################################
##3. Is the sample effort sufficient to reliable capture the structure of the interaction network?
##################################################################################################



```{r}

```



#################################################################################################################
##4.2. Build a dataframe with the ineractions and their weights that you will use yo build your interaction network
#################################################################################################################

#Select the raw data about pair-wise interactions required in the format of "RecruitNet"


```{r}


```

#4.2.1. Obtain the cover data for the canopy species (and Open) species in your study site
#4.2.2. Remove those species lacking cover data 
#4.2.3. Combine data on cover and interaction frequency, and summarize data by canopy-recruit pair across plots.


```{r}

```

#4.2.4.Assess which interactions are significantly different from that expected based on the canopy and open cover 
#4.2.4.1. Retain only facilitation interactions


```{r}


```

#4.2.5. Add indices of interaction strength to those interactions significantly different from the expected values 
#4.2.5.1 and assign 0 to the rest of interactions (no facilitation interactions) and two other weights "Dcr"and "Icr"

```{r}


```

##4.2.6. remove species with non-significant positive interactions

```{r}

```


##################################################################################################
# 5. How does your network looks like?
##################################################################################################

## 5.1. Represent the adjacency matrix using a heatmap (and weight each interaction by the density of recruitment per unit area of canopy species)


```{r fig1, fig.height = 10, fig.asp = 0.9, warning=FALSE, message=FALSE}

```


## 5.2. Represent the network graph using igraph.


```{r fig2, fig.height = 10, fig.asp = 1, warning=FALSE, message=FALSE}

```


#####################################################################################
#6. Questions 
#####################################################################################

###############################
#6.1. From the canopy perspective
###############################

##################################################################################################
#6.1.1. Which species are better nurses (in recruitment enhancement(facilitation) networks?
##################################################################################################

```{r}


```

#####################################################################################################################
#6.1.2. Which nurse species' hold a higher diversity of recruit species (in recruitment enhancement(facilitation) networks?
#####################################################################################################################

```{r}


```

###############################
#6.2.From the recruit perspective
###############################

##################################################################################################
#6.2.1. Which recruit species benefit from more nurses'species (in recruitment enhancement(facilitation) networks?
##################################################################################################

```{r}


```

#####################################################################################
# 8.1. Is there some non-random pattern of interactions in the the observed facilitated and competition networks, such as nestedness?
#(i.e. specialist species tend to interact with generalist species)?
#####################################################################################

```{r}


```
