---
title: "Practice_RN_arid"
output: html_document
date: "2024-09-01"
---

##The aim of the following practice is that you answer some relevant questions that can be aproached using the data provided in RecruitNet to build a Recruitment network, using as an example the study site:Datcamaquis


##########################################################################################################
# 1. Load the functions and that data
##########################################################################################################


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

##############################################################################################################
##2. extract from RecruitNet database the information from the study site:
##############################################################################################################

```{r}


```


##################################################################################################
##3. Is the sample effort sufficient to reliable capture the structure of the interaction network?
##################################################################################################



```{r}

```



#################################################################################################################
##4.1. Build a dataframe with the ineractions and their weights that you will use yo build your interaction network
#################################################################################################################

#Select the raw data about pair-wise interactions required in the format of "RecruitNet"


```{r}


```


##################################################################################################
# 5. How does your network looks like?
##################################################################################################

## 5.1. Represent the adjacency matrix using a heatmap (and weight each interaction by the density of recruitment per unit area of canopy species)


```{r fig1, fig.height = 10, fig.asp = 0.9, warning=FALSE, message=FALSE}

```


## 5.2. Represent the network graph using igraph.


```{r fig2, fig.height = 10, fig.asp = 1, warning=FALSE, message=FALSE}

```




#####################################################################################
#6. Questions 
#####################################################################################


###############################
#6.1.From the canopy perspective
###############################

##################################################################################################
#6.1.1. Which species provide a higher canopy service (i.e. in recruitment networks )?
##################################################################################################

```{r}


```

##################################################################################################
#6.1.2. Which species provide a more diverse canopy service (i.e. in recruitment networks )?
##################################################################################################


```{r}


```


###############################
#6.2.From the recruit perspective
###############################

##################################################################################################
#6.2.1. Which species have a broader recruitment niche (i.e. in recruitment networks )?
##################################################################################################

```{r}


```

############################################################
#7.1. Which species could persist in the community over time?
############################################################

```{r}


```

############################################################
#7.2. Out of all posible interactions, which of them actually occur
#(i.e. does every species recruit under any other species)?
############################################################

```{r}


```
